Species richness destabilizes trophic communities in freshwater ecosystems

### 2020-01-31 manuscript/bef_stability/manuscript.Rmd
Switch to lm4 + bootstrap

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

@@

stab_div_troph <- temporal_network_metrics %>%
  dplyr::select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_bm = log10(biomass_med),
    log_stab = log10(biomass_stab)) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, log_rich, log_stab, log_bm, biomass_med) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

# Compute sem dataset 
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]], 
  hab_press = habitat_pressure, 
  sync = com_data[["sync"]],
  nmds = NULL,
  basin = st_basin 
  )
```

@@

```{r mod-div-total}
library(lme4)
ctrl <- lmeControl(opt='optim')
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optim'))
stab_div_mod <- lmer(data = sem_data,
  #control = ctrl_lmer,
  formula = log_stab ~ log_rich + (1 | basin)
)
stopifnot(!isSingular(stab_div_mod, tol = 1e-05))
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

#
bm_div_mod <- lme(data = sem_data,
  fixed = log_bm ~ log_rich,
  random = ~ 1 + log_rich | basin,
  control = ctrl 
)
bm_div_sum <- piecewiseSEM::rsquared(bm_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

```{r mod-div-troph, fig.show = "hide"}
stab_troph_mod <- lme(data = na.omit(stab_div_troph),
  fixed = log_stab ~ log_rich + troph_group + log_rich:troph_group,
  random = ~ 1 | basin,
  control = ctrl
)

@@

@@

- Main outcome

@@

>>>>

knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.dim = c(7, 7),
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = FALSE
)

@@

stab_div_troph <- temporal_network_metrics_classes %>%
  dplyr::select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_bm = log10(biomass_med),
    log_stab = log10(biomass_stab)) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, log_rich, log_stab, log_bm, biomass_med) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

# Compute sem dataset 
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]], 
  hab_press = habitat_pressure, 
  sync = com_data[["sync"]],
  nmds = NULL,
  basin = st_basin 
  )

@@

```{r mod-div-total}
library(lme4)
ctrl <- lmeControl(opt='optim')
ctrl_lmer <- lmerControl(optCtrl = list(method = 'optimx'))
#If we want that the model converges with basin as random slope,
# we have to remove basin with low data
sem_data %>% group_by(basin) %>% summarise(n())
stab_div_mod <- lmer(
  data = sem_data,
  control = ctrl_lmer,
  formula = log_stab ~ log_rich + (1 | basin)
)

stopifnot(!isSingular(stab_div_mod, tol = 1e-05))
#stab_div_colin <- car::vif(stab_div_mod) not use if one response
stab_div_sum <- piecewiseSEM::rsquared(stab_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

#
bm_div_mod <- lmer(data = sem_data,
  formula = log_bm ~ log_rich + (1 | basin)
)
bm_div_sum <- piecewiseSEM::rsquared(bm_div_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

@@

```{r mod-div-troph, fig.show = "hide"}
stab_troph_mod <- lmer(data = na.omit(stab_div_troph),
  formula = log_stab ~ log_rich + troph_group + log_rich:troph_group + (1 | basin)
  #control = ctrl
)
mySumm2 <- function(.) {
  c(beta=fixef(.),sigma=sigma(.), sig01=sqrt(unlist(VarCorr(.))))
}
stab_troph_boot <- bootMer(stab_troph_mod, mySumm2, nsim = 100)
#plot(test,index=2)
stab_troph_ci <- confint(test)

@@

```{r get-sem-coeff}
coef_stab_sem <- coef(stab_sem) %>%
  group_by(Response) %>%
  nest() %>%
  mutate(sum_c = map_dbl(data, ~ sum(.x$Std.Estimate)))

#cor_mat <- cor(sem_data[, !names(sem_data) %in% c("station", "basin")])
```

@@

- Main outcome:

The effect of richness on biomass stability follows a power law with a negative
coefficient, meaning that it's a decelerating function.

@@

# Appendix

```{r}
mod_list <- list(bm_div_mod, bm_troph_mod, stab_div_mod, stab_troph_mod)
plot_res <- map(mod_list, ~ plot(.x))
plot_grid(plotlist = plot_res)
```
